Symbolic Weighted Language Models and Parsing over Infinite Alphabets

### 2021-05-16 main-V0.tex
split files

@@

%For a complete semiring $\Semiring$, 
%a function $\phi: \Sigma \to \Semiring$, 
%resp. $\eta: \Sigma \times \Delta \to \Semiring$,
%is called

@@

\begin{example}
DTW for sequences of timestamped events **     
\end{example}

@@

\noindent
The above transition functions $\wei_\epsilon$ and $\wei_\Sigma$
can be synthesized as above into a function
$\wei: Q \times (\Sigma \cup \{ \epsilon \}) \times Q \to \Semiring$.

%
%When $\wei_\epsilon(q, q') = \zero$ for all $q, q' \in Q$, 
%the automaton~$A$ is called \emph{without $\epsilon$-transitions}.

@@

\noindent
We define transition functions $\wei'_\epsilon$ and $\wei'_\Sigma$ for $A_{T, s}$, 
together with their synthesized form $\wei'$ as follows: 
for all $1 \leq i, j \leq n$, $q, q' \in Q$, %and $a \in \Sigma$, 
and $b \in \Delta$, 
\[
\begin{array}{lclclcll}
\wei'_\epsilon\bigl(\< i, q>, \< i, q'>\bigr) & = & \wei_\epsilon(q, q') & \mathrm{\ie} &
\wei'\bigl(\< i, q>, \epsilon, \< i, q'>\bigr)   & = & \wei(q, \epsilon, \epsilon, q'), & \\ 
%
\wei'_\epsilon\bigl(\< i, q>, \< i+1, q'>\bigr) & = & \phi(s_i) & 
\multicolumn{5}{l}{\mathrm{where~} \phi = \wei_\Sigma(q, q'),}\\
 & & & \mathrm{\ie} &
\wei'\bigl(\< i, q>, \epsilon, \< i+1, q'>\bigr) & = & \wei(q, s_i, \epsilon, q') & 
\mathrm{if~} i < n,\\ 
%
\wei'_\epsilon\bigl(\< i, q>, \< j, q'>\bigr) & = & \zero & 
%\mathrm{\ie} & \wei'(\< i, q>, \epsilon, \< j, q'>)   & = & \zero & & 
\multicolumn{5}{l}{\mathrm{if~} j \neq i, i+1,}\\ 
%
\wei'_\Sigma\bigl(\< i, q>, \< i, q'>\bigr) & = & \wei_\Delta(q, q') & \mathrm{\ie} &
\wei'\bigl(\< i, q>, b, \< i, q'>\bigr)          & = & \wei(q, \epsilon, b, q'),\\ 
%
\wei'_\Sigma\bigl(\< i, q>, \< i+1, q'>\bigr) & = & \eta_{s_i} &
\multicolumn{5}{l}{\mathrm{where~} \eta = \wei_{\Sigma, \Delta}(q, q'),}\\
 & & & \mathrm{\ie} & \wei'\bigl(\< i, q>, b, \< i+1, q'>\bigr) & = & \wei(q, s_i, b, q') & 
 \mathrm{if~} i < n,\\ 
%
\wei'_\Sigma\bigl(\< i, q>, \< j, q'>\bigr) & = & \zero & \mathrm{\ie} &
\wei'\bigl(\< i, q>, b, \< j, q'>\bigr)     & = & \zero & \mathrm{if~} j \neq i, i+1. 
\end{array}      
\]
We use the assumption that the underlying label theory of $T$ is 
is closed under partial applications (Section~\ref{section:symbols})
in the above construction of $\wei'_\Sigma$.

\noindent
The $\SWA$ wanted is $A_{T, s} = \< Q', \init', \bar{\wei'}, \final' >$.
\qed
\end{proof}

@@

\noindent 
When $\Semiring$ is complete, we call \emph{summary} of a 
to a $\SWT$ $T = \< Q, \init, \bar{\wei}, \final >$
a triplet of functions 
$\mei_\Sigma$, $\mei_\Delta$, and $\mei_{\Sigma, \Delta}$
from $Q \times Q$ into~$\Semiring$, defined by 
\[
\begin{array}{rcll}
\mei_\Sigma(q, q') & = & \bigoplus_{a \in \Sigma} \phi(a) & 
\quad\mathrm{where~} \phi = \wei_\Sigma(q, q') \in \Phi_\Sigma,\\
\mei_\Delta(q, q') & = & \bigoplus_{b \in \Delta} \psi(b) & 
\quad\mathrm{where~} \psi = \wei_\Delta(q, q') \in \Phi_\Delta,\\
\mei_{\Sigma, \Delta}(q, q') & = & 
 \bigoplus_{a \in \Sigma} \bigoplus_{b \in \Delta} \eta(a, b) & 
\quad\mathrm{where~} \eta = \wei_{\Sigma, \Delta}(q, q') \in \Phi_{\Sigma, \Delta}.\\
\end{array}      
\]
Note that these functions are well defined under the assumption that $\Semiring$ is complete.
%A $\SWT$ is called \emph{presented} when these functions, with finite domains, 
%are given explicitly.
For a \SWA, the summary is only defined by $\mei_\Sigma$.

>>>>

% old version

@@

\medskip\noindent
When~$\Semiring$ is complete, 
we call \emph{summary} of a function
$\phi \in \Phi_\Sigma$,
resp. $\eta \in \Phi_{\Sigma, \Delta}$,
the value $\bigoplus_{a \in \Sigma} \phi(a)$, 
resp. $\bigoplus_{a \in \Sigma} \bigoplus_{b \in \Delta} \eta(a, b)$.
Note that, by definition of infinite sums in complete semirings, 
a summary of $\phi \oplus \phi'$ and $\phi \otimes \phi'$ %(and $\alpha \otimes \phi$)
can be computed from summaries of $\phi$ and $\phi'$ in $\Phi_{\Sigma}$, using the operators of $\Semiring$, 
and the same holds for $\eta, \eta' \in \Phi_{\Sigma, \Delta}$. 
%
A label theory is called \emph{effective} when 
summaries of $\phi \oplus \eta$, $\phi \otimes \eta$, $\eta \oplus \psi$, $\eta \otimes \psi$, 
and of the partial applications $\eta_a$ and $\eta_b$, 
can be computed, using the operators of $\Semiring$, 
from summaries of 
$\phi \in \Phi_{\Sigma}$, $\psi \in \Phi_{\Delta}$, $\eta \in \Phi_{\Sigma, \Delta}$.

@@

\begin{example}
Let $\Sigma$ be an alphabet of symbols with timestamps. 
The timestamp of $a \in \Sigma$, denoted by $\mathsf{t}(a)$, is expressed as a rational number.

DTW for sequences of timestamped events **     
\end{example}

@@

\noindent
The above transition functions $\wei_\epsilon$ and $\wei_\Sigma$
can be synthesized as above into a function
$\wei: Q \times (\Sigma \cup \{ \epsilon \}) \times Q \to \Semiring$.
%
%When $\wei_\epsilon(q, q') = \zero$ for all $q, q' \in Q$, 
%the automaton~$A$ is called \emph{without $\epsilon$-transitions}.

A $\SWT$ or $\SWA$ is called \emph{summarized}
when

@@

\noindent
We define transition functions $\wei'_\epsilon$ and $\wei'_\Sigma$ for $A_{T, s}$, 
together with their synthesized form $\wei'$ as follows: 
for all $1 \leq i, j \leq n$, $q, q' \in Q$, %and $a \in \Sigma$, 
and $b \in \Delta$, 
\[
\begin{array}{lclcrcll}
\wei'_\epsilon\bigl(\< i, q>, \< i, q'>\bigr) & = & \wei_\epsilon(q, q') & \mathrm{\ie} &
\wei'\bigl(\< i, q>, \epsilon, \< i, q'>\bigr)   & = & \wei(q, \epsilon, \epsilon, q'), & \\ 
%
\wei'_\epsilon\bigl(\< i, q>, \< i+1, q'>\bigr) & = & \phi(s_i) & 
\multicolumn{5}{l}{\mathrm{where~} \phi = \wei_\Sigma(q, q'),}\\
 & & & \mathrm{\ie} &
\wei'\bigl(\< i, q>, \epsilon, \< i+1, q'>\bigr) & = & \wei(q, s_i, \epsilon, q') & 
\mathrm{if~} i < n,\\ 
%
\wei'_\epsilon\bigl(\< i, q>, \< j, q'>\bigr) & = & \zero & 
%\mathrm{\ie} & \wei'(\< i, q>, \epsilon, \< j, q'>)   & = & \zero & & 
\multicolumn{5}{l}{\mathrm{if~} j \neq i, i+1,}\\ 
%
\wei'_\Sigma\bigl(\< i, q>, \< i, q'>\bigr) & = & \wei_\Delta(q, q') & \mathrm{\ie} &
\wei'\bigl(\< i, q>, b, \< i, q'>\bigr)          & = & \wei(q, \epsilon, b, q'),\\ 
%
\wei'_\Sigma\bigl(\< i, q>, \< i+1, q'>\bigr) & = & \eta_{s_i} &
\multicolumn{5}{l}{\mathrm{where~} \eta = \wei_{\Sigma, \Delta}(q, q'),}\\
 & & & \mathrm{\ie} & \wei'\bigl(\< i, q>, b, \< i+1, q'>\bigr) & = & \wei(q, s_i, b, q') & 
 \mathrm{if~} i < n,\\ 
%
\wei'_\Sigma\bigl(\< i, q>, \< j, q'>\bigr) & = & \zero & \mathrm{\ie} &
\wei'\bigl(\< i, q>, b, \< j, q'>\bigr)     & = & \multicolumn{2}{l}{\zero\ \mathrm{~if~} j \neq i, i+1.}
\end{array}      
\]
We use the assumption that the underlying label theory of $T$ is 
is closed under partial applications (Section~\ref{section:symbols})
in the above construction of $\wei'_\Sigma$.

\noindent
The $\SWA$ wanted is $A_{T, s} = \< Q', \init', {\bar\wei}', \final' >$.
\qed
\end{proof}

@@
